Sketching for Density Estimation

### 2018-11-27 density.tex
asd

\section{Row (or Column) Subset Selection}
Assume you are getting the rows $x_i$  of matrix $X$ in a stream and you want to compute $Z$ such that $\|Z^TZ - X^TX\| \le \eps n$.
For simplicity, assume $x_i$ are all unit length. Moreover, assume each vector in $Z$ must be one of the rows in $X$ up to some constant factor.
This is called the row subset selection problem. 
Since you are trying to minimize the quadratic form, you get that $f(x, q) = \langle x,q \rangle ^2$ and 
$$E(y) = \sum_i s_i \langle x_i,q \rangle ^2 = q^T (\sum_i s_i x_i x_i^T ) q \le \|\sum_i s_i x_i x_i^T\| \le c\sqrt{d}$$

@@

\zk{Selecting $o(1/\eps^2)$ is claimed to be hard by \cite{DBLP:conf/focs/DeshpandeR10} and \cite{DBLP:conf/soda/DeshpandeRVW06}. We need to see if these hold in our setting or only hold for sampling.  Also, \cite{DBLP:conf/soda/GhashamiP14} showed that the impossibility result is limited to frequent directions like settings. Their assumptions don't hold here.}

\section{Streaming Graph Approximation}
Streaming graphs have received a lot of attention in the last several years \cite{}. 
In the standard setting, the edges of a graph $G(V,E)$ are given to an algorithm one ofter the other in arbitrary order.
Here we show how to approximate all quadratic forms $\sum_{(i,j) \in E} (q_i - q_j)^2$ where the values $q$ are not known in advance.
Let $L_e$ be the graph Laplacian of the edge $e$. More accurately, $L_e(i,i) = L_e(j,j) = 1$ and  $L_e(i,j) = L_e(j,i) = -1$ and all other values are zeros.
Then, $\sum_{(i,j) \in E} (q_i - q_j)^2 = q^T \sum_{e \in E}L_e q = q^T L q$. 
In this formulation, this problem reduced to that of row (column) subset selection. 
Note that $L_e = v_ev_e^T$ where $v_e(i) = 1$, $v_e(j)=-1$ and zero else.
Applying the results of the last section we get a streaming mergeable algorithm of approximating graph quadratic forms.
The number of retained edges in the coreset $E'$ is $|E'| = O(\log^{2}(n) \sqrt{n}/\eps)$ and it guaranties that 
$|\sum_{(i,j) \in E} (q_i - q_j)^2  - c \sum_{(i,j) \in E'} (q_i - q_j)^2| \le \eps n$ for all unit norm vectors $q$.

>>>>

\section{Row Subset Selection}
Assume you are getting the rows $x_i$  of matrix $X$ in a stream and you want to compute $Z$ such that $\|Z^TZ - X^TX\| \le \eps n$.
For simplicity, assume $x_i$ are all unit length. Moreover, assume each vector in $Z$ must be one of the rows in $X$ up to some constant factor.
This is called the row subset selection problem. 
Since you are trying to minimize the quadratic form, you get that $f(x, q) = \langle x,q \rangle ^2$ and 
$$E(y) = \sum_i s_i \langle x_i,q \rangle ^2 = q^T (\sum_i s_i x_i x_i^T ) q \le \|\sum_i s_i x_i x_i^T\| \le c\sqrt{d}$$

@@

\el{Selecting $o(1/\eps^2)$ is claimed to be hard by \cite{DBLP:conf/focs/DeshpandeR10} and \cite{DBLP:conf/soda/DeshpandeRVW06}. We need to see if these hold in our setting or only hold for sampling.  Also, \cite{DBLP:conf/soda/GhashamiP14} showed that the impossibility result is limited to frequent directions like settings. Their assumptions don't hold here.} \zk{We can get $1/\eps^2$ though which is better than $\sqrt{d}/\eps$ as long as $\eps > 1/\sqrt{d}$. If $\eps < 1/\sqrt{d}$ then $\sqrt{d}/\eps > d$ - shouldn't it be trivial to get a coreset of size $d$?}

\subsection{optimal design of experiments}
We discuss the problem of linear regression with mean squared error. In the classic setting we are given $n$ points $x_1,\ldots,x_n \in \R^d$ along with $n$ labels $y_1,\ldots,y_n \in \R$ and wish to find $w \in \R^d$ minimizing
$$ \sum_i (w^Tx_i - y_i)^2 $$
In the field of optimal design of experiments, the labels are not given with the data. Instead, the learner must choose in advance a subset of datapoints for which they query the labels. The subset is chosen to provide the least error in a generalization bound, assuming $(x_1,y_1),\ldots,(x_n,y_n)$ are i.i.d from an unknown distribution.

The way this is done in optimal design of experiments is as follows. Let $X$ be the set of $n$ points and consider a subset $B \subseteq X$. The empirical risk minimizer (ERM) for $B$ is 
$$ w_B = \left( \sum_{x \in B} xx^T \right)^{\dagger} y_B $$
%It is known that $w_B$ is an unbiased estimator of $w^*$, the optimal regressor for the distribution, where the noise comes from the realizations of the $y_i$'s. It follows that to minimize 
%$$ E_{x \sim {\cal D}} \ip{w-w^*,x}^2 $$
%which is the variance of $w$ measured by the distribution of data points, it suffice to minimize the second moment
%$$ E_{x \sim {\cal D}} \ip{w,x}^2  \approx w^T \left( \frac{1}{n} \sum_{i=1}^n x_ix_i^T \right) w $$

One can show that for $w_B$,
$$ \sum_i (w_B^Tx_i - y_i)^2 = C \cdot \text{Tr} \left(\left( \sum_{x \in B} xx^T \right)^{\dagger} \right) $$
This only works when $\sum_{x \in B} xx^T$ is full rank, otherwise we need to introduce some regularization and then things get messy.

We can use the results above for optimal design in the case where $w$ is restricted to have a unit Euclidean norm. Something that is commonly practiced. Since we will solve the problem for the subset we choose in an optimal way for gaussian noise, we end up with an unbiased estimator for $w^*$. The error is the variance of that unbiased estimator and minimizing the variance for an unbiased estimator is the same as minimizing its second moment we can aim to minimize
$$ F(w) = w^T \left( \frac{1}{n} \sum_{i=1}^n x_ix_i^T \right) w $$
This is leads to the exact same objective as in column subset selection. Meaning that an $\eps$ core-set for the $x$'s gives a guaranteed error for a subset in the optimal design setting. Specifically we get that we can obtain a core-set of size
$$ \min(1/\eps^2, \sqrt{d}/\eps)$$
For the case of $\eps < 1/\sqrt{d}$, which is an interesting case in linear regression, a random sample of $\sqrt{d}/\eps$ points will not give a generalization bound of $\eps$ (there are known lower bounds, http://jmlr.org/papers/volume16/shamir15a/shamir15a.pdf contradicting this), meaning that we have a provable non-trivial guarantee for active learning. I'm not sure what statements are known already.

In the optimal design setting, they require keeping the covariance matrix. We could come up with a hueristic or even sketching algorithm that avoids maintaining $d^2$ points.
